---
site: "bookdown::bookdown_site"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
  bookdown::pdf_book:
    keep_tex: yes
---

# R syntax

## Assigning values to objects
`<-` is the standard assignment operator. You will assign values to objects (e.g. `x`, `y`, and `j`) using this symbol. For example `x <- 3`  is an object with one element, `x` is assigned a value of 3.

`c` is a common function used for concatenation (think combine): 
```{r, eval = FALSE}
x <- c(3,7,11,19)
y <- c(1,1,1,1)
c(x,y)
y <- c(x,y,5)
```

There are a suite of functions to enter data in the console. The sequence function `seq`:
```{r, eval=FALSE}
y <- seq(from = 1, to = 10)
y <- seq(from = 1, to = 10, by = 2.5)
y <- seq(from = 1, to = 10, length.out = 22)
```

Use the function `rep` (repeat):
```{r, eval=FALSE}
y <- rep(x = 5, times = 4)
x.value <- c(2,3)
rep(x.value, times = 3)
```

## Arithmetic operators
Symbol| Operation
------| ---------
+     | addtion
-     | subtraction
*     | multiplication
/     | division
^     | exponentiation

PEMDAS applies when writing R-code
```{r, eval=FALSE}
# Arithmetic operators in action:
5 + 5 / 2
3 * 2^2
(3*2)^2 
# Arithmetic operators using objects:
z <- 5
w <- c(3,7,9,2)
s <- w[3]
z + s
```

Arithmetic operators practice
Using what you know about parenthesis and PEMDAS, in one line of code do each of the following:  

Number | Excercise
------ | -------------------------------------------------------------------
1.     |Assign the variable `x` to be a vector containing the values 5,5,6,2  
2.     |Assign the variable `y` to be a vector containing the values 3,3,1,7
3.     |Add `x` and `y`
4.     |Substract `y` from `x` 
5.     |Assign `d` as `y` divided by `x`
3.     |Multiply `z` by `s` then add five (`z` and `s` from last practice)
4.     |Add 5 to `z` then multilply by `s`
5.     |Take `z` to the fifth power and then add 2
6.     |Divide `s` by three, then add 33, then take that sum to the 0.5 power

## Dealing with missing values

+ In R, missing values are represented by "NA". 

+ Undefined values (like dividing by zero) are represented by "NaN", not a number. Often missing values are represented with numbers: -1, 99, -9999, etc. This is obviously a problem and should be avoided. You will likely need to use indexing prior using arithmetic operations to replace these values.

The "is.na" function and the "na.rm"" argument:

+ Sometimes we do not know whether there are missing values in our data. 

+ We can use the `is.na` function to test for missing values:

```{r, eval=FALSE}
# is.na tests for missing values
dat.1 <- c(-1,NA,1,1,-1)
is.na(dat.1)
which(is.na(dat.1))
```

We can use the logical `na.rm` argument to remove missing values from our data prior to executing the funciton:
```{r, eval=FALSE}
dat.1 <- c(-1,NA,1,1,-1)
mean(dat.1)
mean(dat.1, na.rm = T)
```

Sometimes we code no data as -1 and that can really screw things up. R does not know that -1 means "no data". However we can replace the -1 with NA. There are many ways to do this, but here is a simple way. 
```{r, eval=FALSE}
# In this example, -1 is coded as a missing data field. Think back to our logical arguments and subsetting exercises.
dat.2 <- c(2,-1,3,4,5)
dat.2 == -1
dat.2[dat.2 == -1]
dat.2[dat.2 == -1] <- NA 
```

```{r, eval=FALSE}
dat.1 <- c(2,NA,3,4,5)
# is.na tests for missing values
is.na(dat.1) 
# returns the element(s) number in the vector that is NA
which(is.na(dat.1))
```
