---
site: "bookdown::bookdown_site"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
  bookdown::pdf_book:
    keep_tex: yes
---

# R syntax 1

## Arithmetic operators
Symbol| Operation
------| ---------
+     | addition
-     | subtraction
*     | multiplication
/     | division
^     | exponentiation

PEMDAS applies when writing R-code
```{r, eval=FALSE}
# Arithmetic operators in action:
5 + 5 / 2
3 * 2^2
(3*2)^2 
# Arithmetic operators using objects:
z <- 5
w <- c(3,7,9,2)
s <- w[3]
z + s
```

## Arithmetic operators practice

Using what you know about parenthesis and PEMDAS, in one line of code do each of the following:  

Number | Excercise
------ | -------------------------------------------------------------------
1.     |Assign the variable `x` to be a vector containing the values 5,5,6,2  
2.     |Assign the variable `y` to be a vector containing the values 3,3,1,7
3.     |Add `x` and `y`
4.     |Substract `y` from `x` 
5.     |Assign `d` as `y` divided by `x`
3.     |Multiply `z` by `s` then add five (`z` and `s` from last practice)
4.     |Add 5 to `z` then multilply by `s`
5.     |Take `z` to the fifth power and then add 2
6.     |Divide `s` by three, then add 33, then take that sum to the 0.5 power

## Missing values (NA)


```{r, eval=FALSE}
# is.na tests for missing values
dat.1 <- c(-1,NA,1,1,-1)
dat.1 + 2
dat.1 + rep(2, length(dat.1))
```

## Dealing with missing values

+ In R, missing values are represented by "NA". 

+ Undefined values (like dividing by zero) are represented by "NaN", not a number. Often missing values are represented with numbers: -1, 99, -9999, etc. This is obviously a problem and should be avoided. You will likely need to use indexing prior using arithmetic operations to replace these values.

The "is.na" function and the "na.rm"" argument:

+ Sometimes we do not know whether there are missing values in our data. 

+ We can use the `is.na` function to test for missing values:

```{r, eval=FALSE}
# is.na tests for missing values
dat.1 <- c(-1,NA,1,1,-1)
is.na(dat.1)
which(is.na(dat.1))
```

We can use the logical `na.rm` argument to remove missing values from our data prior to executing the funciton:
```{r, eval=FALSE}
dat.1 <- c(-1,NA,1,1,-1)
mean(dat.1)
mean(dat.1, na.rm = T)
```

## Dealing with problematic values

Sometimes we code no data as -1 and that can really screw things up. R does not know that -1 means "no data". However we can replace the -1 with NA. There are many ways to do this, but here is a one way. 
```{r, eval=FALSE}
# In this example, -1 is coded as a missing data field. Think back to our logical arguments and subsetting exercises.
dat.2 <- c(2,-1,3,4,5)
dat.2 == -1
dat.2[dat.2 == -1]
dat.2[dat.2 == -1] <- NA 
```

```{r, eval=FALSE}
dat.1 <- c(2,NA,3,4,5)
# is.na tests for missing values
is.na(dat.1) 
# returns the element(s) number in the vector that is NA
which(is.na(dat.1))
```

## Some capabilities of base R statistics

```{r, eval = FALSE}
require(stats)
require(datasets)
require(utils)
```

### Z-scores
```{r, eval = FALSE}
scores <- c(76,80,83,97,100)
?scale
scaled.scores <- scale(scores)
hist(scores)
scores.data.frame <- data.frame(scores,scaled.scores)
scores.data.frame
plot(x = scores.data.frame$scores, 
     y = scores.data.frame$scaled.scores, 
     type = "b",
     xlab = "Raw Test Scores",
     ylab = "Scaled Test Scores")
abline(h = 0)
mean(scores)
abline(v = mean(scores))
```

```{r, eval = FALSE}
z.scores <- c(scores - mean(scores))/sd(scores)
scores.data.frame <- data.frame(scores,scaled.scores,z.scores)
scores.data.frame
```

### Quantiles
Here is an example using the quantile() function:
```{r, eval = FALSE}
?quantile
nile.vect <- Nile
hist(nile.vect, main = "", xlab = "Nile river discharge volume")
quantile(x = nile.vect)
abline(v = quantile(x = nile.vect), col = "red", lwd = 2)
```

### Normality 

Shaprio-Wilk test of normality, the null hypothesis - data are normally distributed

```{r, eval = FALSE}
?shapiro.test
?rbeta
test.dat.1 <- rbeta(n = 500, shape1 = 0.5, shape2 = 5)
hist(x = test.dat.1)
# Test if these data are normally distributed
shapiro.test(x = test.dat.1)
```

```{r, eval = FALSE}
test.dat.2 <- rnorm(n = 500, mean = 0, sd = 1)
hist(x = test.dat.2)
# Test if these data are normally distributed
shapiro.test(x = test.dat.2)
```

### t-test
```{r, eval=FALSE}
head(ToothGrowth,15)
un.dose <- unique(ToothGrowth$dose)
ind.1 <- which(ToothGrowth$dose == un.dose[1])
ind.2 <- which(ToothGrowth$dose == un.dose[2])

t.test(ToothGrowth$len[ind.1], ToothGrowth$len[ind.2])
```

### Linear Models
The lm function can be used to create a simple regression model. 
The `r `lm()` function accepts a number of arguments: 

  + formula: describes the model
    + This is "YVAR ~ XVAR" 
    + YVAR is the dependent (predicted)
    + XVAR is the independent (predictor)

```{r, eval=FALSE}
# Load the UNM_Enroll.csv data file
plot(x = mtcars$wt, 
     y = mtcars$mpg,
     xlab = "Weight", 
     ylab = "mpg", 
     type = "p", 
     las = F)

model.object <- mtcars$mpg ~ mtcars$wt
model.object
lm(model.object)
```

+ The intercept is 37.285 and the coefficient for the unemployment rate is  -5.344  
+ Therefore, the complete regression equation is `r `mpg = 37.285 +  -5.344 x Weight`. 

```{r, eval=FALSE}
# NHST
anova(lm(model.object))
summary(lm(model.object))
pred.y.vals <- predict(lm(model.object))
plot(x = mtcars$wt, 
     y = mtcars$mpg,
     xlab = "Weight", 
     ylab = "mpg", 
     type = "p", 
     las = F)
lines(x = mtcars$wt, pred.y.vals)
```

### Analysis of Variance
```{r, eval=FALSE}
# ANOVAs are linear models with categorical predictors
head(ToothGrowth)
anova.model.object <- ToothGrowth$len ~ ToothGrowth$dose
anova(lm(anova.model.object))
```
