---
site: "bookdown::bookdown_site"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
  bookdown::pdf_book:
    keep_tex: yes
---

# Subsetting and Logical Operators

## Indexing
Indexing is used for "interogation" and "assignment". Interogation means to examine the value of an object. Assigment means to assign the value of an object.
```{r, eval=FALSE}
x <- c(3,7,11,19)
# Assignment
x[c(2,3)] <- c(120,140)
# Interogation
x <- c(3,7,11,19)
x[3]
x[4]
```

Extend the utility of indexing: `[]` by combining with `c`. For example to get the second and third element in the variable "x"
```{r, eval=FALSE}
x <- c(3,7,11,19)
x[1]
x[2]
x[c(1,2)]
x[c(2,1)]
y <- c(3,2,1)
x[y]
```

```{r, eval=FALSE}
# Extend the index using a negative sign
y.vect <- c(1,2,3,4,5,6)
y.vect[-3]
```

Number| Excercise
------| -------------------------------------------------------------------
1.    |Can you get a + to occur on your console?
2.    |How many different ways can you create a vector labled `q` containing two 3's and four 5's? Try some!
3.    |Assign a vector of four elements: 3,7,9 and 2 to `w`.
4.    |Assign the third element of `w` to `s`, where `s` is equal to 6. 
5.    |What is the length of a sequence that starts at 1.1, ends at 9.2, and has increments of 0.894?
6.    |What is the 3rd value of the sequence you created?
7.    |Use the summary function on the sequence object you created, what is returned?



## Logical operators
Symbol | Operation
------ | ---------
<      | less than
>      | greater than
<=     | less than or equal to
>=     | greater than or equal to
==     | exactly equal to

Logical operators are useful in indexing and subsetting data classes. They return `TRUE` and `FALSE` values.
```{r, eval=FALSE}
# logical operators examples
y <- c(2,3,4)
w <- c(3,7,9,2)
y > 1
w == 2
w < 4 | w >= 8
w > 4 & w <= 8
``` 

Combine logical operators with the function `which`
```{r, eval=FALSE}
# logical operators examples
y <- c(2,3,4)
y > 1
which(y > 1)
which(y < 3)
which(y <= 3)
y[which(y <= 3)] <- 0
``` 

Logical operators practice

Number | Excercise
------ | -------------------------------------------------------------------
1.     |Assign a vector `a` to have six elements: 1,0,8,11,12,20
2.     |Which elements in `a` are less than five
3.     |Which elements in `a` are greater than 13 or less than 10



