---
site: "bookdown::bookdown_site"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
  bookdown::pdf_book:
    keep_tex: yes
---

# Assignments

Below you will find:

+ A description of how code will be turned in and evaluated.

+ Four assignments that you will complete during the semester.

+ A description of the Project.

## Turning in code

You will save your code to your individual-linked dropbox account.

Considerations:

1. All files submitted will be .r files and these must be annotated and have descriptive header information. Annotation and descriptive header information is reviewed in the pages of this course's website.

2. For assignments that require you to submit multiple files, all of the files will need to run error free and accomplish the tasks for credit to be earned.

## Assignment 01 - Working in R and RStudio

### Assignment 01.01

+ Make a new script into the 'code' directory, the name of the script will be called 'init.proj.r'

+ This script will be called 'init.proj'

+ The code will populate the working directory with subdirectories: R, data, doc, figs, and output.

+ The goal of the function will be to automatically creates a directory structure outlined in (https://nicercode.github.io/blog/2013-04-05-projects/). i.e. it will have directories: .

### Assignment 01.02

+ Make a new script, the name of the script will be called 'import.and.write.r'

+ This script will do the following:

1. Write the entire mtcars data as a csv file to the 'data' directory (mtcars is in the package 'datasets')

2. Import the the.csv file you wrote, selecting only the first 10 rows only. 

3. From this imported and reduced data frame, make a new data frame that includes a column that is the hp per cylinder (divide hp by cyl)

4. Name this column "hp.per.cyl". 

5. Print the data frame to the screen

6. Save the data frame the 'data' directory as mtcars.red.RData (it will be saved as file type .RData).

### Assignment 01.03

+ Make a new script, the name of the script will be called 'index.practice.r'

1. x <- c("ww", "ee", "ff", "uu", "kk"), use indexing to return:
a. "ee", "ff"
b. "ee"
c. "ff"

2. x <- c("ss", "aa", "ff", "kk", "bb"), use indexing to return: 
a. "uu"
b. NA
c. "kk"

3.  a <- c(2, 4, 6, 8) and b <- c(TRUE, FALSE, TRUE, FALSE)
Write code to determine the maximum and minimum values of a[b].

4. Write an R expression that will return the sum value of 10 for the vector x <- c(2, 1, 4, 2, 1, NA)

5. Write an R expression that will return the sum value of 11 for the vector x <- c(2, 1, 4, 2, 1, NA)

6. Consider the data frame s <- data.frame(first= as.factor(c("x", "y", "a", "b", "x", "z")), second=c(2, 4, 6, 8, 10, 12)). Write two different R statements that will return the output 2, 4, 10, by using the variable first as an index vector.

7. Write three different R statements that will return the positions of 3 and 7 in the vector x <- c(1, 3, 6, 7, 3, 7, 8, 9, 3, 7, 2).

### Assignment 01.04

+ Make a new script, the name of the script will be called 'logical.operators.practice.r'

Using the mtcars data: 

1. Use logical operators to output only those rows of  data  where column  mpg  is between 15 and 20 (excluding 15 and 20).

2. Use logical operators to output only those rows of  data  where column  cyl  is equal to 6 and column  am  is not 0.

3. Output only those rows of  data  where columns  vs  and  am  have the same value 1, solve this without using  ==  operator.

4. Output only those rows of  data  where at least  vs  or  am  have the value 1, solve this without using  ==  or  !=.

### Assignment 01.05

+ Make a new script, the name of the script will be called 'loops.practice.r'

Using the mtcars data: 

1. Write a for loop that iterates over the first seven values of the first numeric column in mtcars and prints the cube of each number using print().

2. Write a for loop that iterates over the first seven values of the first three numeric columns in mtcars and prints values using print().

3. Write a for loop that iterates over the column names of the  iris dataset. The loop will print each column name and the number of characters in the column name in parenthesis. Example output for one iteration of the loop is: "Sepal.Length (12)". Use the following functions print(), paste() and nchar().

4. Write a while loop that prints out standard random normal numbers (use rnorm()) but stops (breaks) if you get a number bigger than 1.

5. Use a while loop to investigate the number of terms required before the product (1)(2)(3)(4)...(n) reaches a value greater than 10 million.

## Assignment 02 - Communicate

```{r, eval=FALSE}
# + Use your function init.proj to populate the project directory.
# 
# + For each of the exercises in: r4ds  '28.2.1 Exercises' (n = 3)
# 
# Wrap each of the exercise codes in a function saved to: 
# 
# 1. function make.fig.01 saved in .r file 'make.fig.01.r'
# 
# 2. function make.fig.02 saved in .r file 'make.fig.02.r'  
# 
# 3. function make.fig.03 saved in .r file 'make.fig.03.r'
# 
# Each of the functions functions will have at least two function arguments that can be changed by the user to alter the aesthetics or data in the figure. The function will write the output figure to a .png file in the 'figs' directory in your project. The title of the figures will be:
# 
# 'fig.01.png', 'fig.02.png', and 'fig.03.png'
```

## Assignment 03 - The Tidyverse and Loops

```{r}

# You will make a project called 'Assignment 03'.
# 
# + Use your function init.proj
# 
# + The goal of this exercise will be to compare the speed of execution of operations on large data frames. We will use with the babynames data from the 'babynames' package (you will need to download this). The babynames dataset has historical information about births of babies in the U.S. View a preview of this dataset by typing babynames in to the console. Yes, this dataset contains 1.8 million observations (that's a lot of data!). 
# 
# + After your inspection of the data, create a file 'speed.test.r', this will be a script, and saved in the correct directory. The .r file will contain two functions:
# 
# + The first function will be called 'tidy.arrange' and will contain code to arrange the babynames tibble in alphabetical order. You will arrange the rows such that the names column is arranged alphabetically. Here you will use tidyverse functionality. The function will have two function arguments, 1. the input data, babynames, and 2, the row numbers that will be sorted. So, for example if all row numbers are sampled, the argument will be 1:dim(babynames)[1]. If you want the first 30 rows to be sorted, the argument will be 1:30 .
# 
# + tidy.arrange will return two objects. The first is a data frame of the ordered babynames, the second is the time that it takes the function to execute. Hint: to get the time use a function that returns the time from computer's internal clock at the start and finish of the function (I generally use package 'matlab' functions 'tic' and 'toc' for this, there are other ways). 
# 
# + Now you will write a script saved as 'arrange.test.r'. This script will run 'tidy.arrange' but the goal will be to sequentially change the second function argument in tidy.arrange to evaluate the time to completion on different sized (number of rows) data.
# 
# + The 'arrange.test.r' script will do the following:
# 
# 1. Source 'tidy.arrange.r'
# 
# 1. Run 'arrange.test' multiple times (at least five). During each of these trials you will:
# 
# a. Change the second function argument of tidy.arrange
# 
# b. Save the output from arrange.test in a data frame. 
# 
# Run 'arrange.test' at least five times and assign these data to an object.
# 
# 2. Using the saved data from the trials you performed: Create a plot that has the number of seconds on the x-axis and the number of rows of the input data on the y-axis. Write this figure to the figs directory as a .png file, 'arrange.test.trial.png'.

```

```{r}
# ## Assignment 04 - Rmarkdown
# 
# You will make a project called 'Assignment 04'.
# 
# Create a Rmarkdown document (.Rmd) that describes and documents the motivation and approach, code and output, and figure(s) you created in Assignment 03 - basically walk an uninformed reader through the assignment. Note when using Rmarkdown, it is easiest to put all of the files (.r, .png, ect.) into a single directory, so no need to make sub-directories in the project. Target: 500 words of text. Cite at least one paper and use the bibtex functionality of Rmarkdown to add a formatted citation (this does not count against your word count).
# 
# This Rmarkdown document 'speed.test.write up.Rmd' will pull the functions created in Assignment 03. Therefore the figure and functions will need to be saved into the Assignment 04 directory. 
```

## Project
### Project Requirement 01: 

You will upload a .Rmd file that I will knit to an .html file. This document will include your name and the title of your project. Note in Rmarkdown doc, it is easiest to put all of the files (.r, .png, ect.) into a single directory, so no need to make sub-directories in the project.

You will write approximately 400 words allocated into three sections:

1.) Introduce your project - three sentences.

2.) Describe the:

+ computational methods and 

+ the data you will use.

This section (the computation section section) will take the most text. You will use topic sentences and paragraphs to describe the general and specific approach you will use to evaluate the data. The data must be larger than 2,000 records and must be publicly available. 

You will keep the scope of the proposal to focus on description of the data and describe in detail how you will manipulate the data with code (i.e. describe the computational methods in detail). The focus of the project is not to review the actual scientific implications (statistical inference and hypothesis testing). Instead, the focus of the project will be data exploration, using Groelmund and Wickham approaches.

3.) Describe the deliverable (at least three) that you will accomplish (one sentence per deliverable),

4.) Present the project idea to the class - 20 minute presentation using only material you have provided me in the .Rmd file.

The document will be titled "Name.Proj.Prop.Rmd", where "Name" is your last name. You will meet in person or by phone, with me, to discuss the project scope and objectives prior to submitting the .Rmd file.   

### Project Requirement 02: 

You will prepare a .Rmd file of your project. I will knit this to a .html file that you will present to the class.

The grade assigned based on the level of complexity of the analysis and amount of original coding. Full credit will be given to projects that showcase your expertise in all aspects of the techniques discussed in this class. Only projects that contain every aspect of import, tidy, transform, visualize, model, and communicate can be given full credit. The code you provide in your project will be extensive and will document your mastery of the tidyverse, data visualization, and original and interesting analysis. My expectation is that the coding will be extensive and demonstrate the philosophy of reproducible workflow that we have studied. The .html knit must knit on my machine.
