---
site: "bookdown::bookdown_site"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
  bookdown::pdf_book:
    keep_tex: yes
---

# R Basics

## Arithmetic operators
Symbol| Operation
------| ---------
+     | addition
-     | subtraction
*     | multiplication
/     | division
^     | exponentiation

PEMDAS applies when writing R-code



```{r, eval=FALSE}
# Arithmetic operators in action:
5 + 5 / 2
3 * 2^2
(3*2)^2 
# Arithmetic operators using objects:
z <- 5
w <- c(3,7,9,2)
s <- w[3]
z + s
```

## Arithmetic operators practice

Using what you know about parenthesis and PEMDAS, in one line of code do each of the following:  

Number | Excercise
------ | -------------------------------------------------------------------
1.     |Assign the variable `x` to be a vector containing the values 5,5,6,2  
2.     |Assign the variable `y` to be a vector containing the values 3,3,1,7
3.     |Add `x` and `y`
4.     |Substract `y` from `x` 
5.     |Assign `d` as `y` divided by `x`
3.     |Multiply `z` by `s` then add five (`z` and `s` from last practice)
4.     |Add 5 to `z` then multilply by `s`
5.     |Take `z` to the fifth power and then add 2
6.     |Divide `s` by three, then add 33, then take that sum to the 0.5 power

## Missing values (NA)

```{r, eval=FALSE}
# is.na tests for missing values
dat.1 <- c(-1,NA,1,1,-1)
dat.1 + 2
dat.1 + rep(2, length(dat.1))
```


## Dealing with missing values

+ In R, missing values are represented by "NA". 

The "is.na" function and the "na.rm"" argument:

+ Sometimes we do not know whether there are missing values in our data. 

+ We can use the `is.na` function to test for missing values:

```{r, eval=FALSE}
# is.na tests for missing values
dat.1 <- c(-1,NA,1,1,-1)
is.na(dat.1)
which(is.na(dat.1))
```

We can use the logical `na.rm` argument to remove missing values from our data prior to executing the funciton:
```{r, eval=FALSE}
dat.1 <- c(-1,NA,1,1,-1)
mean(dat.1)
mean(dat.1, na.rm = T)
```

